---
title: "1. Datos, Modelos y Software. Ejercicios"
date: "Enero 2025, Adair Macías"
output:
  html_document:
    toc: true
    toc_depth: 1
    theme: cerulean
---


<style>
body {
text-align: justify}
</style>

<style type="text/css">
body{/*Normal*/
     font-size: 20px;
}
</style>



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(discrtr) # A companion package for the book Introduction to Discrete Choice Analysis with `R`
library(mlogit) # Multinomial Logit Models
library(readr) # Read Rectangular Text Data
library(plotly)
```



Invoke data set Mode from package mlogit. To do this you need to first load the package. This is a data set with choices about mode of transportation. This is done as follows:

```{r}
data("Mode")
```

Once you have loaded the data set, answer the following questions:
Describe this data set. How many variables are there and of which type (i.e., categorical/quantitative)?

```{r}
glimpse(Mode)
```
*Comentario:*
Hay 9 variables, de las cuales solo una (choice) corresponde al tipo categórico, la demás son de tipo cuantitativo.




How many different modes of transportation are in this data set? What is the most popular mode? What is the least popular mode?

```{r}
summary(Mode)
```
*Comentarios:*
```{r}

modos_transporte <- length(unique(Mode$choice)) #unique para extraer los valores de la columna Choice
print(paste("Los modos de transporte que hay son:", paste(modos_transporte, collapse = ", ")))

unidades <- table(Mode$choice) #table para contar cuantas veces se repite cada modo de transporte

mas_popular <- names(which.max(unidades)) #names para extraer el nombre
print(paste("El modo de transporte más polular es:", paste(mas_popular)))

menos_popular <- names(which.min(unidades))
print(paste("El modo de transporte menos polular es:", paste(menos_popular)))

```


```{r}
ggplot(Mode)+
  geom_bar(aes(choice, fill=choice), color="black", show.legend = FALSE)+
  ggtitle("Distribución de los modos de transporte")+
  theme(plot.title = element_text(hjust = 0.5))
  theme_bw()
```


In general, what is the most expensive mode? The least expensive?


```{r}
ggplot(Mode)+
  geom_density(aes(cost.car, fill="car"), color="black", alpha=0.5)+
  geom_density(aes(cost.carpool , fill="carpool"), color="black", alpha=0.5)+
  geom_density(aes(cost.bus, fill="bus"), color="black", alpha=0.5)+
  geom_density(aes(cost.rail, fill="rail"), color="black", alpha=0.5)+
  scale_fill_manual(values=c("car"= "firebrick", "carpool"="dodgerblue", "bus"="darkgoldenrod2", "rail"="cyan"), name="Mode")+
  xlab("Cost")+
theme_bw()
  
```
*Comentarios:*

```{r}
print("El modo de transporte más caro es: car")
print("El modo de transporte más barato es: carpool")
```



Create a plot showing the univariate distributions of time by car and time by bus. Discuss.

```{r}
grafica_time <- ggplot(Mode)+
  geom_density(aes(time.car, fill="car"), color="black", alpha=0.5)+
  geom_density(aes(time.carpool , fill="carpool"), color="black", alpha=0.5)+
  geom_density(aes(time.bus, fill="bus"), color="black", alpha=0.5)+
  geom_density(aes(time.rail, fill="rail"), color="black", alpha=0.5)+
  scale_fill_manual(values=c("car"= "firebrick", "carpool"="dodgerblue", "bus"="darkgoldenrod2", "rail"="cyan"), name="Mode")+
  xlab("Time")+
theme_bw()


ggplotly(grafica_time)  
```

*Comentarios:*

```{r}
print("Podemos observar que los tiempos entre un modo de transporte y el otro, sean respectivamente car y bus, son muy similares. Pero si observamos el summary del inicio, nos percatamos que el tiempo de bus es mayor (por un par de unidades) al tiempo de car.")
```



How do choices relate to cost by the different modes?

```{r}
varnum <- Mode |> dplyr::select(where(is.numeric))
varnum <- names(varnum)

for (var in varnum) {
grafica_box <- ggplot(Mode)+
  geom_boxplot(aes(choice, eval(as.name(var)), fill=choice), show.legend=FALSE)+
  ylab(var)+
  theme_bw()
  
  print(grafica_box)
    
}



```








