---
title: "Ejercicios Capítulo 5"
format: html
editor: visual
---



```{r}
#| message: false
#| warning: false
#| message: false

library(knitr)
library(tidyverse)
library(discrtr) # A companion package for the book Introduction to Discrete Choice Analysis with `R`
library(dplyr) # A Grammar of Data Manipulation
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(mlogit) # Multinomial Logit Models
library(readr) # Read Rectangular Text Data
#library(stargazer) # Well-Formatted Regression and Summary Statistics Tables
library(gplots) # Various R Programming Tools for Plotting Data
library(ggmosaic) # Mosaic Plots in the 'ggplot2' Framework
library(treemapify)
library(ggridges)
library(ggalluvial)
library(evd)
library(htmlwidgets) # HTML Widgets for R
library(kableExtra) # Construct Complex Table with kable and Pipe Syntax
library(plotly) # Create interactive web graphics
library(tidyr) # Tidy messy data
#library(webshot2) # Take screenshots of web pages
```

```{r setup, include=FALSE}
mc_commute_long <- mc_commute_wide |> 
  mlogit.data(shape = "wide",
              choice = "choice",
              varying = 3:22)
```

Exercise 1

In the example in this chapter we estimated the probabilities of choosing different
modes by sidewalk density setting travel time to the in-sample median. Use
model2 to calculate the probability of choosing different modes by in-sample
median sidewalk density but now for travel times of 20, 30, and 40 min. Discuss
the results.



```{r setup, include=FALSE}
f2 <- mFormula(choice ~ time | sidewalk_density)
model2 <- mlogit(f2, mc_commute_long)
model2 <- mlogit(f2, mc_commute_long, reflevel = "Walk")

mc_commute_predict <- mc_commute_long[1:52, ]
sidewalk_median <- median(mc_commute_predict$sidewalk_density, na.rm=TRUE)
mc_commute_predict$sidewalk_density <- sidewalk_median

# Función para calcular probabilidades para un tiempo dado
calcular_probs_por_tiempo <- function(tiempo) {
  mc_commute_predict$time <- tiempo
  probs <- predict(model2, newdata=mc_commute_predict)
  
  
  probs_df <- as.data.frame(probs)
  probs_df$Tiempo <- tiempo
  
  return(probs_df)
}

probs_20 <- calcular_probs_por_tiempo(20)
probs_30 <- calcular_probs_por_tiempo(30)
probs_40 <- calcular_probs_por_tiempo(40)

#Combino los resultados
resultados_ej1 <- rbind(probs_20, probs_30, probs_40)

kable(resultados_ej1, caption = "Probabilidades por tiempo de viaje con densidad de aceras mediana")
```

Análisis de resultados -esperados-: [Codigo incorrecto(contraintuitivo)]

A medida que aumenta el tiempo de viaje, la probabilidad de caminar disminuye.

Para viajes más largos (40 min), el transporte público (HSR) se vuelve más competitivo.

El coche mantiene una probabilidad relativamente estable, mientras que la bicicleta es menos probable para viajes más largos.



Exercise 2. 
Estimate a model using formula f3 (call it model3). Discuss the output of this
model.

```{r setup, include=FALSE}
f3 <- mFormula(choice ~ 0 | sidewalk_density + time)


model3 <- mlogit(f3, mc_commute_long, reflevel = "Walk")

summary(model3)

```

Análisis de resultados:
Se puede observar que a igualdad de otros factores, caminar es la alternativa más preferida; asi mismo, se puede deducir que los entornos peatonales favorecen modos activos y desincentivan motorizados




Exercise 3
3. Use model3 to calculate the probability of choosing different modes by insample
median sidewalk density but now for travel times of 20, 30, and 40 min.
Discuss the results.


```{r setup, include=FALSE}

calcular_probs_model3 <- function(tiempo) {
  mc_commute_predict$time <- tiempo
  probs <- predict(model3, newdata=mc_commute_predict)
  
  probs_df <- as.data.frame(probs)
  probs_df$Tiempo <- tiempo
  
  return(probs_df)
}

probs3_20 <- calcular_probs_model3(20)
probs3_30 <- calcular_probs_model3(30)
probs3_40 <- calcular_probs_model3(40)

resultados_ej3 <- rbind(probs3_20, probs3_30, probs3_40)
kable(resultados_ej3, caption = "Probabilidades con Modelo 3 por tiempo de viaje")
```
[Codigo incorrecto(contraintuitivo)]



Exercise 4
In the general case, what is the value of the log-likelihood of the null (Equally
Likely) model?

```{r setup, include=FALSE}
n_alternatives <- length(unique(mc_commute_long$alt))
n_choices <- length(unique(mc_commute_long$id))

logLik_null <- n_choices * log(1/n_alternatives)
logLik_null
```



Exercise 5
Use the likelihood ratio test to compare model3 to model2?Discuss the results.
What restrictions would you need to impose in model3 to obtain model2?
```{r setup, include=FALSE}
lrtest(model2, model3)
```


Analisis de resultados:
El test nos dice que el Modelo 3 (con tiempos distintos para cada transporte) sí es mejor que el Modelo 2 (que usa un solo efecto del tiempo para todos). El valor de 0.015 en la prueba -que es menor a 0.05- confirma que esta diferencia es real y no casual.

Para convertir el Modelo 3 en el Modelo 2, se necesitaria:
- Igualar los coeficientes de tiempo (que el efecto de 10 minutos sea el mismo para coche, bici, transporte y caminar), y

- Quitar los efectos específicos de la densidad de aceras en coche/bici/transporte, dejándolo solo para caminar.
